Absolute, relative and rubber Gromov--Witten theory of hypersurfaces in genus one

### 2019-04-16 Reduced Gromov-Witten Theory of Hypersurfaces in Genus One.tex
imported Dhruv changes to introduction and rubber

\begin{customthm}{A}
There exists a log smooth and proper Deligne—Mumford stack giving a modular compactification for the space of genus $1$ maps to projective space with fixed contact order with a hyperplane. 
\end{customthm}

@@

\subsection{Broader contributions} Beyond genus one, we pass through a number of seemingly useful general techniques. First, we explain how to describe degenerate moduli spaces as fiber products. This is substantially more delicate than the standard situation because of the need to work with elliptic singularities and aligned logarithmic structures. This interaction of the degeneration formula with the geometry of curve singularities is a new phenomenon. The results of~\cite{RSPW} suggest the existence of a reduced higher genus Gromov--Witten theory formed by replacing contracted elliptic components with singularities. The discussion here is likely to carry over to this setting.

@@

@@

\begin{cor}
The logarithmically smooth stack $\mathcal{VZ}_{1,\alpha}(\mathbb P^N|H)$ has at worst orbifold singularities, and consequently admits a non-representable cover by a smooth Deligne--Mumford stack.
\end{cor}

@@

\section{Stratification} The logarithmic smoothness established in Theorem~\ref{thm: log-smoothness} implies that $\mathcal{VZ}_{1,\alpha}(\mathbb P^N|H,d)$ is an orbifold toroidal embeddeding. Consequently, the irreducible components of its boundary and their intersections form a stratification of the space. This stratification will be important in the sequel.

@@

\subsection{The rubber space} In our implementation of the relative theory, we will need the naturally ``rubber'' variant of the space of maps $
\mathcal{VZ}_{1,\alpha}(\mathbb P^N|H)$, where the entire curve is mapped into higher levels.

>>>>

\begin{customthm}{A}
There exists a logarithmically smooth and proper Deligne--Mumford stack giving a modular compactification for the space of genus one maps to projective space with fixed contact order with a hyperplane. 
\end{customthm}

@@

\subsection{Broader contributions} Beyond genus one, we pass through a number of seemingly useful general techniques. First, we explain how to describe degenerate moduli spaces as fiber products. This is substantially more delicate than the standard situation because of the need to work with elliptic singularities and aligned logarithmic structures. This interaction of the degeneration formula with the geometry of curve singularities is a new phenomenon. The results of~\cite{RSPW} suggest the existence of a reduced higher genus Gromov--Witten theory formed by replacing contracted elliptic components with singularities\footnote{We have been informed by Jonathan Wise that his Ph.D. candidate Sebastian Bozlee has constructed these higher genus reduced Gromov--Witten invariants. A manuscript is in preparation.}. The discussion here is likely to carry over to this setting.

@@

\subsection*{Acknowledgements} We have benefited from fruitful conversations with friends and colleagues, including Davesh Maulik, Jonathan Wise, and [other people]. Important  aspects of this work were completed while L.B. and N.N. were visiting MIT in Spring 2018 and the University of Cambridge in Winter 2019, and we thank these institutions for their hospitality.

@@

\begin{cor}
The logarithmically smooth stack $\mathcal{VZ}_{1,\alpha}(\mathbb P^N|H)$ has at worst orbifold singularities, i.e. admits a non-representable cover by a smooth Deligne--Mumford stack.
\end{cor}

@@

\section{Stratification}

\noindent The logarithmic smoothness established in Theorem~\ref{thm: log-smoothness} implies that $\mathcal{VZ}_{1,\alpha}(\mathbb P^N|H,d)$ is an orbifold toroidal embeddeding. Consequently, the irreducible components of its boundary and their intersections form a stratification of the space. This stratification will be important in the sequel.

@@

\subsection{Rubber variants} In our implementation of the relative theory, we will need the naturally ``rubber'' variant of the space of maps $
\mathcal{VZ}_{1,\alpha}(\mathbb P^N|H)$, where the entire curve is mapped into higher levels.

Specifically, let $\mathbb P$ denote the projective bundle $\mathbb P(\mathcal O(1)\oplus\mathcal O)$ on $\mathbb P^{N-1}$. Equip this space with the logarithmic structure coming from the $0$ and $\infty$ sections of the bundle. Consider the moduli space $\mathcal{VZ}_{1,\alpha}(\mathbb P/\mathbb G_m)$ of logarithmic maps
\[
C\to \mathbb P[s]\to \mathbb P^{N-1},
\]
which factorize completely, and where automorphisms are considered over the space of maps to $\mathbb P^{N-1}$. Specifically, two maps that differ by a $\mathbb G_m$ translate in the fiber direction are considered equivalent. Total factorization, similar to the previous case, means that the map to the bundle and the map to the base $\mathbb P^{N-1}$ both factorize through possibly different singularities.

\begin{thm}
The stack $\mathcal{VZ}_{1,\alpha}(\mathbb P/\mathbb G_m)$ is logarithmically smooth.
\end{thm}

It is convenient to work with the logarithmic multiplicative group and its torsors. Recall that the \textit{logarithmic multiplicative group} $\mathbb G_{\mathrm{log}}$ is the functor on logarithmic schemes whose value on a logarithmic scheme $S$ is the group of global sections $H^0(S,M_S^{\mathrm{gp}})$. This functor is representable only after a logarithmically \'etale cover.

The bundle $\mathbb P\to \mathbb P^{N-1}$ gives rise to a $\mathbb G_m$-torsors by deleting the zero and infinity sections. Replacing these fibers by their $\mathbb G_{\mathrm{log}}$ compactifications, we obtain a non-representable functor on logarithmic schemes $\mathbb P_{\mathrm{log}}$ and a logarithmically \'etale modification
\[
\mathbb P\to \mathbb P_{\mathrm{log}}.
\]

{\bf Add reference to Wise--R '19 and Marcus--Wise '17 for the multiplicative group.}

\begin{proof}
Consider the stack over logarithmic schemes $\mathcal{VZ}_{1,\alpha}(\mathbb P_{\mathrm{log}})$ of stable logarithmic maps that factorize completely. Here, stability for the map coincides with stability for the projection to $\mathbb P^{N-1}$. The arguments in the previous section show that maps to $\mathbb P$ that factorize completely are unobstructed. Since this space is a logarithmically \'etale cover of $\mathcal{VZ}_{1,\alpha}(\mathbb P_{\mathrm{log}})$, logarithmic smoothness of the latter follows.

The logarithmic multiplicative group $\mathbb G_{\mathrm{log}}$ acts on $\mathcal{VZ}_{1,\alpha}(\mathbb P_{\mathrm{log}})$ by translation. Tautologically, $\mathcal{VZ}_{1,\alpha}(\mathbb P_{\mathrm{log}})$ is a $\mathbb G_{\mathrm{log}}$-torsor over the moduli problem of maps up to this $\mathbb G_{\mathrm{log}}$ translation. It is representable after a logarithmically \'etale modification by a family of nodal curves. It follows that the space  $\mathcal{VZ}_{1,\alpha}(\mathbb P_{\mathrm{log}}/\mathbb G_{\mathrm{log}})$ of maps up to translation that factorize completely is logarithmically smooth.

To conclude the main result, we compare the stacks $\mathcal{VZ}_{1,\alpha}(\mathbb P_{\mathrm{log}}/\mathbb G_{\mathrm{log}})$ and  $\mathcal{VZ}_{1,\alpha}(\mathbb P/\mathbb G_m)$. Since a map to $\mathbb P$ gives rise immediately to a map to $\mathbb P_{\mathrm{log}}$, there is a morphism
\[
\mathcal{VZ}_{1,\alpha}(\mathbb P/\mathbb G_m)\to \mathcal{VZ}_{1,\alpha}(\mathbb P_{\mathrm{log}}/\mathbb G_{\mathrm{log}}).
\]
A straightforward application of the infinitesimal lifting criterion shows that this map is logarithmically \'etale, so the result follows.
\end{proof}

The main point in the proof above is that on the space of non-degenerate maps with smooth domains, the map from $\mathcal{VZ}^\circ_{1,\alpha}(\mathbb P)\to \mathcal{VZ}^\circ_{1,\alpha}(\mathbb P/\mathbb G_m)$ is a $\mathbb G_m$ torsor, and correspondingly on the compact spaces it is a $\mathbb G_{\mathrm{log}}$-torsor. The logarithmically \'etale modification guarantee representability, and turn this into a nodal curve fibration.
